---
title: "Language Autonomy Research EFA - Group A (N=629) - v1"
author: 'by [Charles C.](https://mrcharles.dev)'
output:
  html_document:
    toc: true
    number_sections: false
    toc_depth: 2
    theme: cerulean
    highlight: tango
    df_print: paged
    toc_float:
      collapsed: false
      smooth_scroll: true
  pdf_document:
    toc: true
    toc_depth: '2'
    latex_engine: xelatex
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 6.67, fig.height = 6.67, dpi = 300)
```


# NOTES
```{r notes, echo = FALSE, message = FALSE, warning = FALSE}
cat("
- This is a placeholder
- for the time being.
", "\n")
```


# STEP 01: Load the necessary libraries (auto-install if not available already)
```{r step-01, echo = FALSE, message = FALSE, warning = FALSE}
packages <- c(
  "psych",      # personality, psychometric research
  "corrplot",   # correlation matrix visualization
  "lavaan",     # latent variable modeling
  "semPlot",    # SEM path diagrams
  "readr",      # reading rectangular text data
  "OpenMx",     # extended SEM
  "dplyr",      # a grammar of data manipulation
  "tidyr",      # create tidy data
  "tinytex"     # to compile LaTeX documents
)
# Find missing packages
missing_packages <- packages[!(packages %in% installed.packages()[, "Package"])]
# Install missing packages
if (length(missing_packages) > 0) {
  install.packages(missing_packages, repos = "https://cloud.r-project.org")
}
# Load all required packages
lapply(packages, library, character.only = TRUE)
```


# STEP 02: Load dataset into a dataframe
```{r step-02, echo = TRUE, message = FALSE, warning = FALSE}
dataset01 <- read.csv("lar_group_A_629_v0.csv")
names(dataset01) # see the variable names
```


# STEP 03: Select only the necessary continuous variables
```{r step-03, echo = TRUE, message = FALSE, warning = FALSE}
dataset02 <- dataset01 %>%
  select(num_range("item_", 1:89, width = 2)) %>%  # item_01 ... item_89
  select(where(is.numeric))                         # keep continuous only
```


# STEP 04: Assumption check - remove rows with missing values
```{r step-04, echo = TRUE, message = FALSE, warning = FALSE}
# rows with any NA (to inspect)
rows_with_na <- dataset02 %>% filter(if_any(everything(), is.na))
# drop them
dataset02_clean <- dataset02 %>% drop_na()
# quick report
removed <- nrow(dataset02) - nrow(dataset02_clean)
cat("Removed", removed, "rows (", round(100*removed/nrow(dataset02), 1), "%).\n")
```


# STEP 05. Assumption check - flag high-correlation item pairs only (above .80)
```{r step-05, echo = TRUE, message = FALSE, warning = FALSE}
# ---- tweakable settings ----
r_cutoff <- 0.80   # show pairs with |r| >= this
p_cutoff <- 0.05   # optionally filter by p-value (set to 1 to ignore)
adjust_method <- "holm"  # "none", "holm", "bonferroni", "BH", etc.

# ---- correlation test ----
ct <- psych::corr.test(dataset02_clean, use = "pairwise", adjust = adjust_method)
R  <- ct$r
P  <- ct$p

# ---- build tidy upper-triangle table ----
idx <- which(upper.tri(R), arr.ind = TRUE)

tab_high <- tibble(
  item1 = colnames(R)[idx[, 1]],
  item2 = colnames(R)[idx[, 2]],
  r     = R[upper.tri(R)],
  p     = P[upper.tri(R)]
) %>%
  mutate(abs_r = abs(r)) %>%
  filter(abs_r >= r_cutoff, p <= p_cutoff) %>%
  arrange(desc(abs_r)) %>%
  select(item1, item2, r, p)

# ---- print a compact report ----
cat("High-correlation pairs (|r| >=", r_cutoff, "), p <=", p_cutoff,
    " (p adjusted:", adjust_method, ")\n\n")

if (nrow(tab_high) == 0) {
  cat("No pairs met the criteria.\n")
} else {
  print(tab_high, n = Inf)
}
```


# STEP 06. Bartlett's test of sphericity
```{r step-06, echo = TRUE, message = FALSE, warning = FALSE}
cortest.bartlett(dataset02_clean)
```


# STEP 07. KMO test
```{r step-07, echo = TRUE, message = FALSE, warning = FALSE}
KMO(dataset02_clean)
```


# STEP 08. Determine number of factors using eigenvalues from correlation test and scree plot of eigenvalues
```{r step-08, echo = TRUE, message = FALSE, warning = FALSE}
dataset02_clean_ev <- eigen(cor(dataset02_clean))$values
plot(dataset02_clean_ev, type = "b", main = "Eigenvalues (Scree Test)", xlab = "Factor Number", ylab = "Eigenvalue")
```


# STEP 09. Run parallel analysis to determine number of factors from dataset directly
```{r step-09, echo = TRUE, message = FALSE, warning = FALSE}
fa.parallel(dataset02_clean, fa = "fa", n.iter = 100, show.legend = TRUE, main = "Scree Plot & Parallel Analysis")
```

# STEP 10. EFA with oblimin rotation and using 6 factors
```{r step-10, echo = TRUE, message = FALSE, warning = FALSE}
dataset02_clean_efa_oblimin_rotation <- fa(dataset02_clean, nfactors = 6, rotate = "oblimin", fm = "ml")
print(dataset02_clean_efa_oblimin_rotation$loadings, cutoff = 0.1)
print(dataset02_clean_efa_oblimin_rotation$loadings, cutoff = 0.4)
```


# STEP 11. EFA with oblimin rotation and using 7 factors
```{r step-11, echo = TRUE, message = FALSE, warning = FALSE}
dataset02_clean_efa_oblimin_rotation <- fa(dataset02_clean, nfactors = 7, rotate = "oblimin", fm = "ml")
print(dataset02_clean_efa_oblimin_rotation$loadings, cutoff = 0.1)
print(dataset02_clean_efa_oblimin_rotation$loadings, cutoff = 0.4)
```


# STEP 12. Confirm 7-factor structure, identify strong anchor indicators, spot problematic cross-loaders, weak items with low h^2 (communality) and high u^2 (uniqueness)
```{r step-12, echo = TRUE, message = FALSE, warning = FALSE}
efa7 <- fa(dataset02_clean, nfactors = 7, rotate = "oblimin", fm = "ml") 

# diagnostic print 
print(efa7, cut = 0.30, sort = TRUE) # shows which items are cross-loaders

# keep these too 
efa7$Phi # factor correlations - what factors are too strongly correlated with each other (>.8)
efa7$communality # h2 values - how much an item’s variance is explained by the extracted factors (low h2 means noise, e.g. <.2)
```


# STEP 13. Visualise factor loadings using pre-determined EFA model
```{r step-13, echo = TRUE, message = FALSE, warning = FALSE}
fa.diagram(dataset02_clean_efa_oblimin_rotation, cut=0.2, digits = 3, main = "Factor Loadings Diagram")
```


# STEP 14. BONUS - confirm findings above using lavaaan
```{r step-14, echo = TRUE, message = FALSE, warning = FALSE}
dataset02_clean_efa_lavaan <- efa(data = dataset02_clean [,1:87], nfactors = 1:7)
summary(dataset02_clean_efa_lavaan)
```
